Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-02-06 manuscript/manuscript.Rmd
Added code for the robustness experiment.

# Generate table of asymmetric generalised normal distribution parameters.
n_distributions <- 100

@@

# Optimise k1, k2 so that sum of absolute differences with target-lambda is
# minimised for each combination of source and weighting method.

>>>>

# Generate table of asymmetric generalised normal distribution parameters.
n_distributions <- 10

@@

# Optimise k1, k2 so that sum of absolute differences with target-lambda is
# minimised for each combination of source and weighting method.
experiment_args <- coro::generator(
  function(){

    # Non-robust transformation.
    fun_args <- list(
      "method" = "yeo_johnson",
      "robust" = FALSE,
      "shift" = TRUE)

    opt_args <- list()

    yield(list(
      "name" = "non-robust",
      "fun_args" = fun_args,
      "opt_args" = opt_args
    ))

    for(robustness_source in c("original", "transformed", "residual")){

      def_opt_limits <- list("k1"=c(0.0, 10.0), "k2"=c(0.0, 10.0))

      if(robustness_source == "original"){
        def_opt_limits$k1[2] <- 1.0
        def_opt_limits$k2[2] <- 1.0

        def_opt_init <- list("k1" = 0.8, "k2" = 0.95)

      } else if(robustness_source == "transformed"){
        def_opt_init <- list("k1" = 1.5, "k2" = 3.0)

      } else if(robustness_source == "residual"){
        def_opt_init <- list("k1" = 0.15, "k2" = 1.0)
      }

      for(robustness_weighting_function in c("step", "triangle", "cosine")){

        fun_args <- list(
          "method" = "yeo_johnson",
          "robust" = TRUE,
          "shift" = TRUE,
          "weight_method" = paste0(robustness_source, "_", robustness_weighting_function),
          "backup_use_default" = FALSE)

        opt_limits <- def_opt_limits
        opt_init <- def_opt_init

        # Step-weighting only has k1 as a parameter.
        if(robustness_weighting_function == "step"){
          opt_limits$k2 <- NULL
          opt_init$k2 <- NULL
        }

        yield(list(
          "name" = paste0(robustness_source, "-", robustness_weighting_function),
          "fun_args" = fun_args,
          "opt_args" = list("initial" = opt_init, "limits" = opt_limits)
        ))
      }
    }
  })

experiments <- coro::collect(experiment_args())

# Function for computing transformer parameters under optimisation constraints.
compute_robust_lambda <- function(
    x,
    fun_args,
    opt_args){

  # Custom parser.
  ..outlier_parser <- function(
    x,
    fun_args,
    opt_args
  ){
    # Create transformer.
    transformer <- suppressWarnings(do.call(
      power.transform::find_transformation_parameters,
      args=c(
        list("x" = x$x),
        fun_args,
        opt_args)))

    parameter_data <- c(
      x$parameters,
      opt_args,
      list(
        "weight_method" = fun_args$weight_method,
        "lambda" = transformer@lambda,
        "shift" = transformer@shift)
    )

    return(parameter_data)
  }

  parameter_data <- lapply(
    x,
    ..outlier_parser,
    fun_args = fun_args,
    opt_args = opt_args)

  return(parameter_data)
}

# Inner optimisation function that sets the loss.
..optimisation_function <- function(
  opt_param,
  x,
  fun_args){

  # Parse options.
  opt_args <- list()
  if(length(opt_param) >= 1) opt_args <- c(opt_args, list("k1"=opt_param[1]))
  if(length(opt_param) >= 2) opt_args <- c(opt_args, list("k2"=opt_param[2]))

  parameter_data <- compute_robust_lambda(
    x = x,
    fun_args = fun_args,
    opt_args = opt_args)

  lambda_error <- sapply(
    parameter_data,
    function(x) (abs(x$lambda - x$target_lambda)))

  return(sum(lambda_error))
}

# Outer optimisation function
.optimisation_function <- function(
    experiment,
    x){

  # Run optimiser.
  if(length(experiment$opt_args) > 0){

    x0 <- unlist(experiment$opt_args$initial)

    lower <- experiment$opt_args$limits$k1[1]
    upper <- experiment$opt_args$limits$k1[2]

    if(length(experiment$opt_args$limits) > 1){
      lower <- c(lower, experiment$opt_args$limits$k2[1])
      upper <- c(upper, experiment$opt_args$limits$k2[2])
    }

    h <- nloptr::sbplx(
      x0 = x0,
      fn = ..optimisation_function,
      lower = lower,
      upper = upper,
      control=list("xtol_rel"=1e-3, "ftol_rel"=1e-4),
      x = x,
      fun_args = experiment$fun_args)

    if(length(h$par) == 1){
      return(list("name"=experiment$name, "k1" = h$par, "value" = h$value))

    } else {
      return(list("name"=experiment$name, "k1" = h$par[1], "k2" = h$par[2], "value" = h$value))
    }

  } else {
    value <- ..optimisation_function(
      opt_param=list(),
      x = x,
      fun_args=experiment$fun_args)

    return(list("name"=experiment$name, "value" = value))
  }
}

lapply(
  experiments,
  .optimisation_function,
  x=x
)
